---
title: "Prediction Assignment Write Up"
author: "John Mastapeter"
date: "11/18/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

## Prelimary Steps

In order to preform any kind of predictive analysis, the necessary R libraies must be loaded and the online data downloaded and scraped of uncessary data prior to the making the calculations.

## Libraries

```{r necessary libraries}
#Please download the following libraries to complete the code;
library(RCurl)
library(knitr)
library(caret)
library(rattle)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
```

## Set Working Directory and Download Data

Set pre-established working directory to save data.

```{r setwd and download}
#Set working directory
setwd("C:/Users/mastapeterj/Documents/Coursera_DataScience/PredictionAssignmentWriteUp")
#download files
#pml testing
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "pmltraining.csv", method = "libcurl")
#pml training
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "pmltesting.csv", method = "libcurl")
#Read data into R
trainingpml <- read.csv("pmltraining.csv")
testingpml <- read.csv("pmltesting.csv")
#Review Data
training_head <-head(trainingpml)
testing_head <-head(testingpml)
#If interested or necessary, print review of data
#training_head
#testing_head
```

##Cleaning the Data
```{r clean data}
#set seed
set.seed(12345)
#Set up initial training partition to use during analysis
inTrain <- createDataPartition(trainingpml$classe, p = 0.7, list = FALSE)
train_set <- trainingpml[inTrain, ]
test_set <- trainingpml[-inTrain, ]
dim(train_set)
dim(test_set)

#Remove all Near Zero Values whicch will not affect the analysis
nzvs <- nearZeroVar(train_set)
train_set <- train_set[, -nzvs]
test_set <- test_set[, -nzvs]
dim(train_set)
dim(test_set)

#Remoove all NA values that will not affect the analysis
allna <- sapply(train_set, function(x) mean(is.na(x))) > 0.95
train_set <- train_set[, allna==FALSE]
test_set <- test_set[, allna==FALSE]
dim(train_set)
dim(test_set)

#Remove the first through fifth columns, as they add no valuable information for analysis
train_set <- train_set[, -(1:5)]
test_set <- test_set[, -(1:5)]
dim(train_set)
dim(test_set)
```

##Determine Correlation through Fixed Point Clustering
```{r FPC Plot}
cor_Matrix <- cor(train_set[,-54])
corrplot(cor_Matrix, order = "FPC", method = "color", type = "lower", tl.cex = 0.8, tl.col=rgb(0 ,0 ,0))
```

##Determine correlation through Random Forests
```{r Ranndom Forest Plot}
set.seed(12345)
control_rf <- trainControl(method = "cv", number = 3, verboseIter = FALSE)
modfit_randforest <- train(classe ~., data=train_set, method = "rf", trControl = control_rf)
modfit_randforest$finalModel

predict_randforest <- predict(modfit_randforest, newdata = test_set)
confmat_randforest <- confusionMatrix(predict_randforest, test_set$classe)
confmat_randforest
plot(confmat_randforest$table, col= confmat_randforest$byClass, main = paste("Random Forest - Accuracy=", round(confmat_randforest$overall['Accuracy'], 4)))
```

##Determine correlation through Decision Tree
```{r Decision Tree Plot}
set.seed(12345)
modfit_dectree <- rpart(classe~., data = train_set, method = "class")
fancyRpartPlot((modfit_dectree))

predict_dectree <-predict(modfit_dectree, newdata = test_set, type = "class")
confmat_dectree <- confusionMatrix(predict_dectree, test_set$classe)
confmat_dectree

plot(confmat_dectree$table, col = confmat_dectree$byClass, main = paste("Decision Tree - Accuracy = ", round(confmat_dectree$overall['Accuracy'], 4)))
```

##Determine correlation through Generalilzed Boosted Model
```{r Generalied Boosted Model}
set.seed(12345)
control_gbm <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
modfit_gbm <- train(classe~., data = train_set, method = "gbm", trControl = control_gbm, verbose = FALSE)
modfit_gbm$finalModel

predict_gbm <- predict(modfit_gbm, newdata = test_set)
confmat_gbm <- confusionMatrix(predict_gbm, test_set$classe)
confmat_gbm

plot(confmat_gbm$table, confmat_gbm$byClass, main = paste("GBM - Accuracy = ", round(confmat_gbm$overall['Acccuracy'], 4)))
```